---
title: "Lab_9"
author: "Gabrielle"
date: '2022-11-21'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, results=FALSE)
library(tidyverse)
library(here)
library(ggpubr) # for some graphic applications that extend ggplot2
library(janitor)
library(broom) # used to make tables
library(knitr) # used to make table
library(lubridate)
```

Start by importing this IDH data. There are two variables. The variable per_disturbance is a manipulated percent of disturbance of a landscape. The species_richness variable is the number of unique species found at each manipulated disturbance patch.

```{r}
IDH <- read_csv(here("data","IDH.csv"))

birds_2021 <- read_csv(here("data","Sedgwick_2021_ebird_obs_mod.csv"))

```

Let us plot the raw data:

```{r, results=TRUE}
scatter<- ggplot(IDH, aes(x=per_disturbance, y=species_richness)) +geom_point() +
  labs(x="Disturbance (%)", y="Species richness") + xlim(0,100) +ylim(0,25)+ theme_bw()

scatter
```


### Models
While we know the data are not linear, we may consider some linear models and some non-linear polynomial models.

m_0: There is no relationship between species richness and disturbance (mean only model)

m_1: There is a linear relationship between species richness and disturbance (simple linear regression). We know from visual inspection this is not a good model.$y=b_0 + b_1 x$

m_2a: a 2nd order polynomial with 
intercept at 0.  $y=b_1 x + b_2 x^2$

m_2b: a 2nd order polynomial with estimated intercept. $y=b_0 + b_1 x + b_2 x^2$

m_3: a 3rd order polynomial with estimated intercept. $y=b_0 + b_1 x + b_2 x^2 + b_3 x^3$


Let us program a suite of models to consider.



#### models of polynomial fit
```{r}
m_0 <- lm(species_richness ~ 1, data=IDH) # mean only model

m_1 <- lm(species_richness ~ per_disturbance, data=IDH) # Simple linear regression

m_2a <- lm(species_richness ~ -1 + per_disturbance + I(per_disturbance^2), data=IDH) # 2nd degree polynomial no constant (intercept)
## I = "as is" command
## run a command without actually changing or adding a variable to the data set

m_2a <- lm(species_richness ~ -1 + per_disturbance + I(per_disturbance^2), data=IDH)

m_2b <- lm(species_richness ~ per_disturbance + I(per_disturbance^2), data=IDH)

m_3 <- lm(species_richness ~ per_disturbance + I(per_disturbance^2) + I(per_disturbance^3), data=IDH)


```

- inspect the coefficients 
- apply same inspection tools used for linear regression to the polynomial regression

This makes a nice table of the model name, followed by useful statistics for model selection
```{r}
BIC_list <- c(BIC(m_0), BIC(m_1), BIC(m_2a), BIC(m_2b), BIC(m_3))

model_output <-rbind(data.frame(glance(m_0)),data.frame(glance(m_1)),data.frame(glance(m_2a)),data.frame(glance(m_2b)), data.frame(glance(m_3))) %>% select(BIC) 

model_output <- mutate(model_output, delta.BIC = BIC-min(BIC_list))
model_output$model<-c("Model 0", "Model 1","Model 2a","Model 2b", "Model 3")
model_output<-model_output[,c("model", "BIC", "delta.BIC" )]


kable(model_output, format = "markdown", digits = 3, caption = "BIC, and Delta.BIC for the IDF models. Delta BIC > 7 indicates models that should be dismissed from further consideration.")
```
- the lowest BIC of the list is the BEST
  - MODEL 2a is the best
  


Graphic of data and model
```{r, results = TRUE}
scatter_m2a <- ggplot(IDH, aes(x = per_disturbance, y = species_richness)) + 
  geom_point() +
  labs(x = "Disturbance (%)", y = "Species richness") +
  xlim(0,100) +
  ylim (-5,25) +
  theme_minimal() +
  stat_smooth(method = "lm", formula = y ~ -1 + x + I(x^2), se = FALSE)

# the -1 is saying to drop the intercept
# this is the 

scatter_m2a
```

